---
title: "R Notebook"
output: html_notebook
---
# Initializations
## 1.1 Required Libraries
```{r}
library(ggplot2)
library(data.table)
options(warn=-1)
```

## 1.2 Load the Data
```{r}
marker_dt <- as.data.table( read.delim("marker.txt"))
growth_dt <- as.data.table( read.delim("growth.txt"))
gene_dt <- as.data.table( read.delim("gene.txt"))
genotype_dt <- as.data.table( read.delim("genotype.txt"))
expression_dt <- as.data.table( read.delim("expression.txt"))
```

# 2. Tidy The Data

### Expressions 
First the expressions data is melted. The strain and environment values on the column names will be values under the columns 'Strain' and 'Medium', repectively.
```{r}
exp_melted <- melt(expression_dt, 
                   id.vars=c(), 
                   measure.vars = colnames(expression_dt), 
                   variable.name = "Strain", 
                   value.name = "Value")

exp_melted[,Medium := Strain]
exp_melted[,Medium:=gsub(pattern = "\\..*$",replacement = "",x=Medium)]
exp_melted[,Strain:=gsub(pattern = "^.*\\.",replacement = "",x=Strain)]
exp_melted[, id:= 1:length(Strain)]
setcolorder(exp_melted, c("id","Strain", "Medium", "Value"))
head(exp_melted)
```

### Genotype

Genotype is melted so that the marker variables will not be in the columns but under the "Marker" column. And the values column is named as "Source", which indicates the source of the corresponding marker.

```{r}
genotype_melted <- melt(genotype_dt,
                        id.vars = c("strain"), 
                        measure.vars = setdiff(colnames(genotype_dt), c("strain")),
                        variable.name = "Marker", value.name = "Source")

head(genotype_melted)
```

## Growth Differences in Different Media
```{r}
growth_melted <- melt(growth_dt, id.vars = "strain", variable.name = "medium", value.name = "growth")

ggplot(growth_melted, aes(medium, growth)) + geom_violin(na.rm = T)

```


## Plot Describing Expression Vs. Growth for Each Strain
```{r}
exp_grouped<- exp_melted[, mean := mean(Value), by=.(Medium, Strain)]
growth_vs_exp <- merge(exp_grouped,growth_melted, by.x = c("Medium", "Strain"), by.y =c("medium","strain") )

ggplot(growth_vs_exp, aes(mean, growth)) + geom_point() + facet_wrap(~Medium)

```

## The Best Growing Strains For Each Media
```{r}
growth_melted[,mean:=mean(growth), by=strain]
growth_melted[order(mean,decreasing = T)]
```



